Spin canonical divisors of log stacky curves

### 2015-06-29 stacky_write_up/spin_cring.tex

%\begin{cor}
%Let $D$ be a fractional $\frac{a}{b}$-canonical divisor of a genus 0 stacky curve with denominator $b > 2$ when $\frac{a}{b}$ is written in reduced form.
%For a similar reason as in corollary ~\ref{cor:genus_0_positive_delta}, $D$ must be equivalent to an effective divisor (in this case even when $\delta= 0$), so by corollary ~\ref{cor:effective_Q_divisor_can_ring} we get the desired result.
%\todo{Move this Lemma:}
%\begin{lem}\label{lem:add_point_effective_divisor_no_deg_1_generators}
%Let $(X,\Delta,\halfcan)\to (X,\Delta',\halfcan)$ be the natural map of genus $g$ log curves with $\Delta=\Delta'+P$, such that $2\halfcan\sim K_X+\Delta$, $2\halfcan'\sim K_X+\Delta'$, and $\Delta=\Delta'+P$ with $\Delta'\ne 0$ for some $P\not\in supp(\halfcan)$. 
%Suppose $\deg(\lfloor{\halfcan}\rfloor)=\deg(\lfloor \halfcan'\rfloor)\ge 0$, so $R_{\halfcan}$ has a
%generator $u$ in degree 1, and $R_{\halfcan'}$ has no new generators in degree 0. Suppose $D$ is generated by $u, x_1, \ldots, x_m$ in degree at most $\bar{d}$ with relations generated 
%in degree at most $\bar{\tau}$.
%\begin{itemize}
%\item $R_{D'}$ is generated over $R_D$ by elements $y_1, y_2, y_3$ in degrees 2, 2, and 3 respectively. 
%\item If $I$ and $I'$ are the ideal of relations of $R_D$ and $R_{D'}$ respectively, then 
	%\initial_\prec(I') &= \initial_\prec(I) k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] \\
										 %&+ \langle y_i x_j \rangle \\
										 %&+ \langle y_3^2, y_3y_2, y_2y_1, y_1^2\rangle \\
%\item A Gr\"{o}bner basis with initial terms described above minimally generates $I'$ over $I$.
%\end{itemize}
%Similar to all the other Lemmas\todo{combine sections?}

@@

\begin{enumerate}
	\item In the case $g \geq 2$, Proposition ~\ref{prop:relation_11} gives a bound on the degree of generators of the relations of a log spin canonical ring $R_L$. Additionally, by Remark ~\ref{rem:relations_generation_ten}, any curve of genus 2 with half canonical divisor $L = P$ has spin canonical ring $R_L$ with relations in degree 10. This raises the question of whether the bound of 11 on the degree of the relations in $R_L$ can be reduced to $10$. In particlar, we make the following conjecture.
\begin{conjecture}
\label{conj:relations-10}
If $X$ is a curve of genus $g \geq 2$ and $L \in \di X$ with $2L \sim K,$ the ideal of relations in $R_L$ is generated in degree at most $10$. More precisely, if $x_1,\ldots, x_n$ is a minimal set of generators for $R_L$ and $\phi:k[y_1,\ldots, y_n] \rightarrow R_L, y_i \mapsto x_i,$ then $I_L := \ker \phi$, is generated in degree at most 10. 
\end{conjecture}
Note that in the case $\dim L \geq 2$ and $L$ is basepoint-free, by Neves ~\cite[Proposition III.12]{neves:halfcan}, $I_L$ is generated in degree at most 8. Additionally, if Conjecture ~\ref{conj:relations-10} were true, by Theorem ~\ref{thm:main} we would immediately obtain the following corollary, providing a more concise statement than that of Theorem ~\ref{thm:main}.
\begin{cor}
\label{cor:main-refined}
Let $(\sx,\Delta,L)$ be a log spin curve over a perfect field $k$ with signature $(g;e_1,\ldots, e_r;\delta)$ and let $R_L$ be the log canonical ring. Suppose Conjecture ~\ref{conj:relations-10}. Then, $R_L$ is generated in degree at most $e = \max(5,e_1,\ldots, e_r)$ with relations in degree at most $2e,$ apart from a finite list of exceptions.
\end{cor}
	\item Another direction for further research is to extend the results of this paper to fractional modular forms, just as this paper extends the results of Voight and Zureick-Brown \cite{vzb:stacky} from even weight modular forms to odd weight modular forms. In particular, we pose the following question:
\begin{question}
\label{ques:fractional-weight}
If $\sx$ is a stacky curve and $D \in \di \sx$ with $nD \sim K,$ where $K$ is the canonical divisor of $\sx$, can one bound the degrees of generators and relations of $R_D$?
\end{question}
Note that that the above question is answered affirmatively in the case that $\sx$ has signature $(0;e_1,\ldots, e_r;\delta)$ in \todo{Aren't we supposed to mention fractional weight modular forms after an early inductive lemma? Can we include this as a remark somewhere?} \todo{Include reference to the fractional weight modular forms remark, in this remark, we should probably reference \cite[Remark 10.2.3]{vzb:stacky}}. In fact, we show that if $D \in \di \sx$ so that $nD = K$ for some $n \geq 2$, then $R_D$ is generated in degree at most $e=\max(5,e_1,\ldots, e_r)$ with relations in degree at most $2e$. Additionally, Lemma ~\ref{lem:raise-stacky-order} naturally extends to the setting fractional weight modular forms, as noted in Remark ~\ref{rem:raise-fractional-order}.
	\item In Voight and Zureick Brown \cite[Definition 2.2.7]{vzb:stacky}, the notion of a generic initial ideal is defined, which encapsulates the idea of whether the relations for $R_L$ are generically chosen, in a precise sense. The proof of Theorem ~\ref{thm:main} decidedly does not yield a generic initial ideal, as generators and relations are constructed in a highly nongeneric manner. In particular, Lemma ~\ref{lem:raise-stacky-order} constructs generators with non-maximal pole orders at certain points, making the relations non-generic. 
\begin{question}
\label{ques:generic-initial}
If $(\sx,\Delta,L)$ is a log spin curve with signature $\sigma$, does there exist a function $\sigma \mapsto d(\sigma)$ so that the generic initial ideal {\rm(} with some specified ordering{\rm)} is generated in degree at most $d(\sigma)$? Can one write down the generic initial ideal explicitly?
\end{question}
	\item As noted in Remark ~\ref{rem:explicit-generators}, the proof of Theorem ~\ref{thm:main} given in this paper gives an explicit procedure for computing the generators and relations of $R_L$ when the genus of $\sx$ is $0$ or $1$.
\begin{question}
\label{ques:algorithm}
Can one give an algorithm for explicitly computing the generators and relations for an arbitrary genus $g \geq 2$ curve?
\end{question}
Suppose the generators and relations of $R_{L_X}$ are known, where $(\sx, \Delta, L)$ is a spin canonical ring, $X$ is the coarse space of $\sx$, and $L_X \leq L$ is a half canonical divisor of $X$. In this case our proof of Theorem ~\ref{thm:main} yields an algorithm for determining the generators of $L$. This raises the following Petri-like question:
\begin{question}
\label{ques:half-canonical-petri}
Given a {\rm(}non-stacky {\rm)} curve $X$ and a half canonical divisor $L \in \di X$, can one write down a set of minimal generators and minimal relations for $R_L$? If so, can one give the generic ideal of $R_L$ or the generic initial ideal of $R_L$, as defined in Voight and Zureick Brown \cite[Definition 2.2.7]{vzb:stacky}?
\end{question}
In particular, if Question ~\ref{ques:half-canonical-petri} were answered affirmatively one would be able to write down explicit generators and relations for $R_L$ on a stacky curve, essentially by inductively applying Lemma ~\ref{lem:raise-stacky-order}.
	\todo{This last one doesn't seem particularly interesting. Should we remove it?}
	\item While Theorem ~\ref{thm:main} gives a minimal set of generators for the log spin canonical ring $R_L,$ the set of relations is not minimal. In many of the $g = 0$ and $g=1$ cases, it is not too difficult to see that our inductive procedure also yields a minimal set of relations for $R_L$. One might investigate whether the generators and relations given by the inductive proof of Theorem ~\ref{thm:main} are always minimal.
	\end{enumerate}

>>>>

%\todo{Move this Lemma:}
%\begin{lem}\label{lem:add_point_effective_divisor_no_deg_1_generators}
%Let $(X,\Delta,\halfcan)\to (X,\Delta',\halfcan)$ be the natural map of genus $g$ log curves with $\Delta=\Delta'+P$, such that $2\halfcan\sim K_X+\Delta$, $2\halfcan'\sim K_X+\Delta'$, and $\Delta=\Delta'+P$ with $\Delta'\ne 0$ for some $P\not\in supp(\halfcan)$. 
%Suppose $\deg(\lfloor{\halfcan}\rfloor)=\deg(\lfloor \halfcan'\rfloor)\ge 0$, so $R_{\halfcan}$ has a
%generator $u$ in degree 1, and $R_{\halfcan'}$ has no new generators in degree 0. Suppose $D$ is generated by $u, x_1, \ldots, x_m$ in degree at most $\bar{d}$ with relations generated 
%in degree at most $\bar{\tau}$.
%Then 
%\begin{itemize}
%\item $R_{D'}$ is generated over $R_D$ by elements $y_1, y_2, y_3$ in degrees 2, 2, and 3 respectively. 
%\item If $I$ and $I'$ are the ideal of relations of $R_D$ and $R_{D'}$ respectively, then 
%
%\begin{align*}
	%\initial_\prec(I') &= \initial_\prec(I) k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] \\
										 %&+ \langle y_i x_j \rangle \\
										 %&+ \langle y_3^2, y_3y_2, y_2y_1, y_1^2\rangle \\
%\end{align*}
%\item A Gr\"{o}bner basis with initial terms described above minimally generates $I'$ over $I$.
%\end{itemize}
%\end{lem}
%\begin{proof}
%Similar to all the other Lemmas\todo{combine sections?}
%\end{proof}

@@

\begin{enumerate}
	\item In the case $g \geq 2$, Proposition ~\ref{prop:relation_11} gives a bound on the degree of generators of the relations of a log spin canonical ring $R_L$. Additionally, by Remark ~\ref{rem:relations_generation_ten}, any curve of genus 2 with half canonical divisor $L = P$ has spin canonical ring $R_L$ with relations in degree 10. This raises the question of whether the bound of 11 on the degree of the relations in $R_L$ can be reduced to $10$. In particlar, we make the following conjecture.
\begin{conjecture}
\label{conj:relations-10}
If $X$ is a curve of genus $g \geq 2$ and $L \in \di X$ with $2L \sim K,$ the ideal of relations in $R_L$ is generated in degree at most $10$. More precisely, if $x_1,\ldots, x_n$ is a minimal set of generators for $R_L$ and $\phi:k[y_1,\ldots, y_n] \rightarrow R_L, y_i \mapsto x_i,$ then $I_L := \ker \phi$, is generated in degree at most 10. 
\end{conjecture}
Note that in the case $\dim L \geq 2$ and $L$ is basepoint-free, by Neves ~\cite[Proposition III.12]{neves:halfcan}, $I_L$ is generated in degree at most 8. Additionally, if Conjecture ~\ref{conj:relations-10} were true, by Theorem ~\ref{thm:main} we would immediately obtain the following corollary, providing a more concise statement than that of Theorem ~\ref{thm:main}.
\begin{cor}
\label{cor:main-refined}
Let $(\sx,\Delta,L)$ be a log spin curve over a perfect field $k$ with signature $(g;e_1,\ldots, e_r;\delta)$ and let $R_L$ be the log canonical ring. Suppose Conjecture ~\ref{conj:relations-10}. Then, $R_L$ is generated in degree at most $e = \max(5,e_1,\ldots, e_r)$ with relations in degree at most $2e,$ apart from a finite list of exceptions.
\end{cor}
	\item Another direction for further research is to extend the results of this paper to divisors $D \in \di \sx$ on a stacky curve $\sx,$ where $nD \sim K.$ The canonical rings of such divisors would translate to rings of fractional weight modular forms. As defined in Adler and Ramanan, immediately following \cite[Corollary 24.5]{adler:moduli}, a {\bf modular form of fractional weight} $\frac{2a}{n}$ on $X(p)$ is a section of a line bundle $\eta$ so that there exists a line bundle $\gamma$ with $\gamma^{\otimes a} \cong \eta$ and $\gamma^{n} \cong \Omega_{X(p)}^1$. Adler and Ramanan \cite[Appendix 2]{adler:moduli} construct an explicit basis of weight $\frac{4}{5}$ modular forms for $X(11)$. Fractional modular forms also make an apperance in Milnor \cite[$\mathsection$ 6]{milnor:fractional-weight}. In this work, Milnor shows that for a triangle group $\Gamma$ of signature $(0; e_1,e_2,e_3; 0)$ there exist generators $f_1,f_2,f_3$ of a certain ring of fractional weight modular forms, satisfying $f_1^{e_1}+f_2^{e_2}+f_3^{e_3} = 0$.

Having motivated the importance of fractional weight modular forms, we pose the following question:

\begin{question}
\label{ques:fractional-weight}
If $\sx$ is a stacky curve and $D \in \di \sx$ with $nD \sim K,$ where $K$ is the canonical divisor of $\sx$, can one bound the degrees of generators and relations of $R_D$?
\end{question}
Note that that the above question is answered affirmatively in the case that $\sx$ has genus $0$ and $D$ is effective, as follows by inductively applying Lemma ~\ref{lem:sat-1}: If $\sx$ has signature $(g;e_1, \ldots, e_r; \delta)$ then $D$ is generated in degree at most $e = \max(e_1, \ldots, e_r)$ with relations in degree at most $2e$. Additionally, it may be possible to modify the proof of Lemma ~\ref{lem:raise-stacky-order} to extend to the setting of fractional weight modular forms. Given suitable generalizations of these lemmas, one may be able to approach the problem of finding generators and relations for fractional weight modular forms using techniques similar to those in this paper.
	\item In Voight and Zureick Brown \cite[Definition 2.2.7]{vzb:stacky}, the notion of a generic initial ideal is defined, which encapsulates the idea of whether the relations for $R_L$ are generically chosen, in a precise sense. The proof of Theorem ~\ref{thm:main} decidedly does not yield a generic initial ideal, as generators and relations are constructed in a highly nongeneric manner. In particular, Lemma ~\ref{lem:raise-stacky-order} constructs generators with non-maximal pole orders at certain points, making the relations non-generic. 
\begin{question}
\label{ques:generic-initial}
If $(\sx,\Delta,L)$ is a log spin curve with signature $\sigma$, does there exist a function $\sigma \mapsto d(\sigma)$ so that the generic initial ideal {\rm(} with some specified ordering{\rm)} is generated in degree at most $d(\sigma)$? Can one write down the generic initial ideal explicitly?
\end{question}
	\item As noted in Remark ~\ref{rem:explicit-generators}, the proof of Theorem ~\ref{thm:main} given in this paper gives an explicit procedure for computing the generators and relations of $R_L$ when the genus of $\sx$ is $0$ or $1$.
\begin{question}
\label{ques:algorithm}
Can one give an algorithm for explicitly computing the generators and relations for an arbitrary genus $g \geq 2$ curve?
\end{question}
Suppose the generators and relations of $R_{L_X}$ are known, where $(\sx, \Delta, L)$ is a spin canonical ring, $X$ is the coarse space of $\sx$, and $L_X \leq L$ is a half canonical divisor of $X$. In this case our proof of Theorem ~\ref{thm:main} yields an algorithm for determining the generators of $L$. This raises the following Petri-like question:
\begin{question}
\label{ques:half-canonical-petri}
Given a {\rm(}non-stacky {\rm)} curve $X$ and a half canonical divisor $L \in \di X$, can one write down a set of minimal generators and minimal relations for $R_L$? If so, can one give the generic ideal of $R_L$ or the generic initial ideal of $R_L$, as defined in Voight and Zureick Brown \cite[Definition 2.2.7]{vzb:stacky}?
\end{question}
In particular, if Question ~\ref{ques:half-canonical-petri} were answered affirmatively one would be able to write down explicit generators and relations for $R_L$ on a stacky curve, essentially by inductively applying Lemma ~\ref{lem:raise-stacky-order}.
	\todo{This last one doesn't seem particularly interesting. Should we remove it?}
	\item While Theorem ~\ref{thm:main} gives a minimal set of generators for the log spin canonical ring $R_L,$ the set of relations is not minimal. In many of the $g = 0$ and $g=1$ cases, it is not too difficult to see that our inductive procedure also yields a minimal set of relations for $R_L$. One might investigate whether the generators and relations given by the inductive proof of Theorem ~\ref{thm:main} are always minimal.
	\end{enumerate}
